\documentclass[itdr]{subfiles}


\newcommand{\travelunit}[1]{\FPeval{\result}{trunc(#1 * 5 : 0)}\result}

\begin{document}

\chapterx{Appendix A: Additional and Alternative Rules}
\index{Additional Rules}

``\title'' is intended as a rules-light game. Keep this in mind when deciding to use any of the rules presented in this appendix.

\section{Characters}

\subsection{Epic Characters}

If you want player characters to be more powerful, roll 2d6~+~6 for Ability Scores, and d4~+~2 for HP.

\vfill
\break

\section{Magic}
\index{Magic}

\subsection{Magic Mishaps}
\index{Magic Mishaps}

Improper use of a magical device, spellcasting inside an anti-magic zone, destruction of a magic item, or any other dangerous interaction with magic might lead to a mishap. Succeed on a \save{WIL} or suffer a random consequence.

\begin{dtable}[cX]
	\textbf{d100} & \textbf{Mishap} \\
	1--4	&	You exude strong smell for a day. (d4) 1:~mint, 2:~garlic, 3:~vinegar, 4:~sulphur.	\\
	5--8	&	Your clothes' colour changes randomly.	\\
	9--12	&	Your clothes grow one size. Gain Disadvantage on \saves{DEX} until this problem is amended.	\\
	13--16	&	Your eye colour changes randomly.	\\
	17--20	&	Your eyes shed bright light for a day.	\\
	21--24	&	Your hair colour changes to a random one (new hair grows normal).	\\
	25--28	&	Your hair falls out.	\\
	29--32	&	Your hair grows to a yearly amount.	\\
	33--36	&	Your skin turns a saturated shade of a random colour for d12 months.	\\
	37--40	&	Your skin is covered in a random growth for d12 months. (d4) 1:~fur, 2:~scales, 3:~feathers, 4:~spines.	\\
	41--43	&	You disappear for a minute.	\\
	44--46	&	You are stunned until Rest.	\\
	47--49	&	You fall unconscious until Rest.	\\
	50--52	&	You are invisible for an hour or until you attack or cast a spell.	\\
	53--55	&	Your ears become pointed and hairy.	\\
	56--58	&	You are deafened until Rest.	\\
	59--61	&	Your voice is very loud until Rest.	\\
	62--64	&	You are mute until Rest.	\\
	65--67	&	You see invisible things for an hour.	\\
	68--70	&	You are blinded until Rest.	\\
	71--72	&	You are obscured by a could of smoke.	\\
	73--74	&	Your HP drop to 0.	\\
	75--76	&	Your HP are restored.	\\
	77--78	&	You double in size for an hour. Gain Advantage on \saves{STR} and increase your weapon Damage dice by one (up to d12).	\\
\end{dtable}
\begin{dtable}[cX]
\textbf{d100} & \textbf{Mishap} \\
	79--80	&	You halve in size for an hour. Gain Disadvantage on \saves{STR} and lower your weapon Damage dice by one (down to d4).	\\
	81--82	&	Your main weapon shrinks to one-sixteenth its size for an hour.	\\
	83--84	&	Your tongue becomes forked.	\\
	85--86	&	Your canines grow long and sharp.	\\
	87--88	&	You grow a tail.	\\
	89--90	&	You grow horns.	\\
	91	&	You grow gills.	\\
	92	&	Your nails grow into sharp talons (bonus d6 unarmed Damage die).	\\
	93	&	Your teeth fall out.	\\
	94	&	Your skin becomes very tough. Gain Armour~1 when not wearing any Armour.	\\
	95	&	One of your Ability Scores increases by one (up to 20). (d6) 1--2:~STR, 3--4:~DEX, 5--6:~WIL.	\\
	96	&	One of your Ability Scores decreases by one (down to 3). (d6) 1--2:~STR, 3--4:~DEX, 5--6:~WIL.	\\
	97	&	You grow a random body part.	\\
	98	&	You lose a random body part.	\\
	99	&	Your clothes burst in flames. Take d6 damage now and d6 at the end of your next turn unless extinguished.	\\
	100	&	You are petrified.	\\
\end{dtable}

\vfill

\header{Random Color}
\begin{dtable}[cXcX]
	\textbf{d12} & \textbf{Color} & \textbf{d12} & \textbf{Color} \\
	1 & snow white		& 7	& lemon yellow \\
	2 & ash grey		& 8	& malachite green \\
	3 & jet black		& 9 & sky blue \\
	4 & crimson red		& 10& ultramarine blue \\
	5 & chestnut brown	& 11& lavender violet \\
	6 & pumpkin orange	& 12& orchid magenta \\
\end{dtable}

\header{Random Body Part}
\begin{dtable}[cXcX]
	\textbf{d20} & \textbf{Part} & \textbf{d20} & \textbf{Part} \\
	1--4	& \multicolumn{3}{l}{a body part created by a magic mishap} \\
	5--7	& finger	& 14--16& foot \\
	8--10	& hand		& 17--18& lower leg \\
	11--12	& forearm	& 19	& leg \\
	13		& arm		& 20	& eye \\
\end{dtable}

\break

\subsection{Speed Casting}
\index{Speed Casting}

Use this option in place of Prepared Spells rule.

Your focus allows you to cast any Spell you know in a number of combat rounds equal to its Circle~+~1. To speed it up, take 2hp Damage ignoring Armour per round of reduced cast time. This Damage occurs when the spell is cast, and at 0hp it targets WIL instead of STR.


\section{Resources}
\index{Resources}

\begin{dtable}[cXcX]
	\textbf{Amount} & \textbf{Description} & \textbf{Average} & \textbf{Price} \\
	1 & running out	& 1		& $\times$~1 \\
	2 & low			& 2		& $\times$~d6 \\
	3 & enough		& 4		& $\times$~2d6 \\
	4 & plenty		& 7		& $\times$~3d6 \\
	5 & excess		& 13	& $\times$~4d6 \\
\end{dtable}

Each time you spend a resource, roll a d6. If you roll over the Amount, decrease it by one. On zero the resource is depleted.

If you search some place for the resource, roll a d6. If you roll over the Amount, increase it by one to a maximum of 5.

When buying resources to increase the Amount by one (up to 5), pay its price multiplied by your current Amount~x~d6.


\section{Siege Engines}
\index{Siege Engines}
\index{Damage}

Siege engines require a crew of two and a whole turn to reload.

\begin{dtable}[lclX]
	\textbf{Engine} & \textbf{Cost} & \textbf{Damage} & \textbf{Ammo} \\
	Ballista 	& 1g 	& d12 		& 10s bolt \\
				& 	 	& d10 		& 5s ball  \\
	\rowcolor{dColor1}\multirow{-2}{*}{Catapult} & \multirow{-2}{*}{1g}	& d10 Blast & 20s bomb \\
	\rowcolor{dColor2}Cannon & 2g	& d12 Blast & 25s shot \\
\end{dtable}

The weight of a siege engine is about 1 ton.

\vfill
\break

\cleartoleftpage
\section{Travel}
\index{Travel}

Travel for 8 hours/day before resting for the night.

\index{Grid}
\subparagraph{Grid} of \travelunit{1}-mile tiles simplifies distance calculations.

\index{Terrain}
\index{Speed}
\begin{dtable}[lXcc]
\textbf{Terrain} & \textbf{Example} & \textbf{Miles} & \textbf{Grid} \\
Clear		& grassland, road, trail 			& \travelunit{5} & 5 \\
Rugged		& desert, forest, hills, mud, snow	& \travelunit{4} & 4 \\
Difficult 	& jungle, mountains, river, swamp	& \travelunit{3} & 3 \\
\end{dtable}

To speed up calculations chose a dominant terrain for each day of travel and apply it to the whole day. 

\subparagraph{Concurrent Activities} like exploring, sneaking, foraging, etc. reduce the speed by \travelunit{2} miles/day.

\index{Encumbrance}
\subparagraph{Encumbrance} while travelling on foot should be 50lb or less, otherwise halve your travelling speed.

\subparagraph{Going Astray} is a possibility when traversing unfamiliar or heavily obscured terrain, in a dense fog or heavy rain, etc. If you have some additional advantage in navigation, roll a d6, otherwise roll a d4:

\begin{dtable}[cX]
	\textbf{d6} & \textbf{Outcome} \\
	1	& \textbf{Lost!} You wander off to an unknown place. \\
	2	& \textbf{Going in circles.} No travel progress today. \\
	3	& \textbf{Meandering.} Halved travelled distance. \\
	4--6& \textbf{On course.} \\
\end{dtable}	

\index{Food}
\index{Forage}
\subparagraph{Food} consumption should be at least 2lb of food and 4lb (roughly \sfrac{1}{2}gal) of water daily, and twice as much water in hot climate (except camels). Horses and such require $\times$10, and elephants $\times$100 as much forage and water. A day without enough water or a week without enough food results in d4 STR Loss.

\index{Forced March}
\subparagraph{Forced March} doubles the speed. Rest for the next day, or, if using a mount, get a fresh one instead.

\index{Rest}
\subparagraph{Rest} for a day for every 6 days travelled. Not resting, or marching more than 8 hours per day results in d4 STR Loss.

\index{Large Groups}
\subparagraph{Large Groups} move at --\travelunit{1} miles/day.

\index{Mounts}
\subparagraph{Mounts} get +\travelunit{1} miles/day on Clear terrain, and --\travelunit{1} miles/day on Difficult terrain. Camels don't suffer such penalty in desert, and elephants --- in jungle.

\index{Mule}
\index{Donkey}
\index{Horse}
\index{Camel}
\index{Elephant}
\begin{dtable}[Xccl]
	\textbf{Mount}  & \textbf{Cargo} & \textbf{Riders} & \textbf{Cost} \\
	Mule, Donkey	& \sfrac{1}{5}t (400lb)	& 1	& 5s \\
	Horse, Camel	& \sfrac{1}{4}t (500lb)	& 2 & 1g \\
	Elephant		& 2t (4000lb)	& 8	& 5g \\
\end{dtable}

\index{Vehicle}
\subparagraph{Vehicles} get +\travelunit{1} miles/day on roads, --\travelunit{1} on Rugged terrain, and can't pass through Difficult terrain.

\index{Cart}
\index{Carriage}
\index{Wagon}
\begin{dtable}[Xcccl]
\textbf{Vehicle} & \textbf{Horses} & \textbf{Cargo} & \textbf{Passengers} & \textbf{Cost} \\
Cart	& 1	& \sfrac{1}{2}t	& 4	& 30s \\
Carriage& 2 & 1t			& 8	& 60s \\
Wagon	& 4	& 2t			& 16& 1g \\
\end{dtable}


\index{Cargo}
\index{Riders}
\index{Passengers}
\subparagraph{Passengers} occupy \sfrac{1}{8}t of cargo space. Cargo and Riders/Passengers values are mutually exclusive.


\vfill


\subsection{Waterborne Travel}
\index{Travel!waterborne}
\index{Waterborne Travel|see {Travel, waterborne}}
\index{Boats}
\index{Ships}

Travel for 12 hours/day. With a double crew, you can switch shifts to continue travelling at night.

\index{Vehicle!waterborne}
\index{Rowboat}
\index{Sailboat}
\index{Keelboat}
\index{Sailing Ship}
\index{Longship}
\index{Galley}
\begin{dtable}[lYYYYl]
\textbf{Vehicle} & \textbf{Miles} & \textbf{Grid} & \textbf{Crew} & \textbf{Cargo} & \textbf{Cost} \\
Raft (100 ft\textsuperscript{2})& \travelunit{2}&2 & 1& \sfrac{1}{4}t & --- \\
Rowboat		& \travelunit{3}	& 3	& 1		& 1t	& 50s \\
Sailboat	& \travelunit{12}	& 12& 1		& 5t	& 15g \\
Keelboat	& \travelunit{6}	& 6	& 10	& 20t	& 25g \\
Longship	& \travelunit{18}	& 18& 50	& 10t	& 100g \\
Sailing ship& \travelunit{18}	& 18& 10	& 100t	& 150g \\
Galley		& \travelunit{18}	& 18& 100	& 150t	& 200g \\
\end{dtable}

Keelboats, Longships and Galleys have both sails and oars, but can't go against the wind under sail.

\subparagraph{Covered Distance} depends on weather and other conditions. Going upstream reduces covered distance by \travelunit{2} miles/day, and going downstream increases it by the same amount. Makeshift rafts move downstream only, with a speed of the stream.

\index{Fare}
\subparagraph{Fare} might vary from 1p per person to cross a river or lake, to 1s per person for each 5 miles travelled in a long-distance voyage.

\index{Passengers}
\subparagraph{Accommodated passengers} occupy 1t of cargo space plus \sfrac{1}{10}t (200lb) of cargo space per month of travel for food and water storage.

\index{Siege Engines}
\subparagraph{Siege Engines} could be mounted on keelboats (1), sailing ships (2), and galleys (3).


\break


\subsection{Weather}
\index{Weather}
\index{Wind|see {Weather}}

Keep in mind that different climates might require adjusting the tables.

\header{Temperature}
\begin{dtable}[cX]
	\textbf{d6} & \textbf{Temperature} \\
	1	& colder than usual \\
	2--5& normal \\
	6	& warmer than usual \\
\end{dtable}

\header{Sky}
\begin{dtable}[cX|cl]
	\textbf{d20} & \textbf{Sky} & \textbf{d20} & \textbf{Sky} \\
	1--4 & clear	& 13--14 & drizzle or fog \\
	5--8 & cloudy	& 15--18 & rain or snow \\
	9--12& overcast & 19--20 & storm or snowstorm \\
\end{dtable}

\header{Wind Direction}
\begin{dtable}[cX]
	\textbf{d8} & \textbf{Direction} \\
	1--3& adverse \\
	4--5& side \\
	6--8& favourable \\
\end{dtable}
When following prevailing wind's direction, roll 2d8 and take the higher result; when going against it --- take the lower one.

\subparagraph{Wind Force} might affect your sailing speed.

\begin{dtable}[cXcc]
	~ & \textbf{Wind} & \multicolumn{2}{c}{\textbf{Sailing Multiplier}} \\
	\textbf{d20} & \textbf{Force} & \textbf{Adverse or Side} & \textbf{Favourable} \\
	1--2	& calm		& $\times$0 & $\times$0 \\
	3--6	& breeze	& $\times$\sfrac{1}{3}	& $\times$\sfrac{1}{2} \\
	7--14	& average	& $\times$\sfrac{1}{2}	& $\times$1 \\
	15--18	& strong	& $\times$\sfrac{2}{3}	& $\times$1\sfrac{1}{2} \\
	19--20	& gale		& $\times$0	& $\times$2 \\
\end{dtable}

Ships exposed to gale in open sea roll for a gale damage each 6 hours.

\header{Gale Damage}
\begin{dtable}[cX]
	\textbf{d8} & \textbf{Damage} \\
	1 	& \textbf{Wrecked.} Ship, cargo, and \sfrac{1}{2} crew is lost. \\
	2 	& \textbf{Broken mast.} No sailing speed. \\
	3 	& \textbf{Broken half of oars.} \sfrac{1}{2} rowing speed. \\
	4 	& \textbf{Torn sail.} \sfrac{1}{2} sailing speed. \\
	5--6& \textbf{Overboard.} Lost d6 crew members. \\
	7--8& \textbf{All is fine.} \\
\end{dtable}

\vfill
\break

\subsection{Aerial Travel}
\index{Travel!aerial}
\index{Aerial Travel|see {Travel, aerial}}

Flying creatures travel for 8 hours/day before resting for the night. Flying magic items have energy to function for the same daily amount of time.

\begin{dtable}[Xlccc]
	\textbf{Mount} & \textbf{Example} & \textbf{Miles} & \textbf{Grid} & \textbf{Riders} \\
	Small				& pixie		& 40  & 8  &---\\
	Medium 				& harpy		& 40  & 8  & 1 \\
	Large				& griffon	& 80  & 16 & 2 \\
	Large, fast			& pegasus	& 120 & 24 & 2 \\
	Huge				& dragon	& 80  & 16 & 8 \\
	Magic \mbox{device}	& broom		& 80  & 16 & 1 \\
	Magic \mbox{vehicle}& carpet	& 40  & 8  & 8 \\
\end{dtable}

Full speed is only possible with \sfrac{1}{2}~of riders or less. Otherwise the speed is halved.

\index{Vehicle!aerial}
\subparagraph{Aerial vehicles} travel for 12 hours/day. Double crew allows to continue travelling at night.

\index{Balloon}
\index{Airship}
\begin{dtable}[lccccl]
	\textbf{Vehicle} & \textbf{Miles} & \textbf{Grid} & \textbf{Crew} & \textbf{Cargo} & \textbf{Cost} \\
	Balloon	& 40 & 8 & 1 & 1t & 25g \\
	Airship	& 40 & 8 & 10& 10t& 200g \\
\end{dtable}

Balloons and airships are affected by the winds in a same manner as sailing ships.

\subparagraph{Balloons} always follow the wind direction. Each 3 hours of travel you may change altitude to catch a preferable wind (roll for a new wind direction).


\begin{dbox}
	\index{Units of Measure}
	\index{Distance|see {Units of Measure}}
	\index{Volume|see {Units of Measure}}
	\index{Weight|see {Units of Measure}}
	\paragraph{Units of Measure}
	
	\subparagraph{Distance}
	\begin{itemize}
		\item 1 mile is 1760 yards or 5280 feet
		\item 1 yard is 3 feet or 36 inches
		\item 1 foot is 12 inches
	\end{itemize}
	
	\subparagraph{Volume}
	\begin{itemize}
		\item 1 gallon is 4 quarts or 8 pints
		\item 1 quart is 2 pints or 32 ounces
		\item 1 pint is 16 ounces
	\end{itemize}
	
	\subparagraph{Weight}
	\begin{itemize}
		\item 1 ton is 2000 pounds
		\item 1 pound is 16 ounces
	\end{itemize}
\end{dbox}

\break

\section{Random Tables and Inspiration}

\subsection{Random Characters}
\index{Random Characters}

Random tables for quick character generation.

\paragraph{Abilities}

\begin{dtable}[c|YYYY]
	\textbf{d20} & \textbf{STR} & \textbf{DEX} & \textbf{WIL} & \textbf{Money} \\
	1	& 16& 12& 10& 8s \\
	2	& 10& 12& 16& 8s \\
	3	& 12& 10& 15& 8s \\
	4	& 15& 12& 10& 8s \\
	5	& 10& 14& 12& 8s \\
	6	& 12& 10& 14& 8s \\
	7	& 14& 10& 12& 7s \\
	8	& 12& 15& 10& 6s \\
	9	& 10& 11& 13& 8s \\
	10	& 14& 12& 9 & 7s \\
	11	& 11& 14& 10& 7s \\
	12	& 8 & 16& 13& 5s \\
	13	& 11& 13& 9 & 8s \\
	14	& 9 & 11& 14& 7s \\
	15	& 13& 11& 9 & 7s \\
	16	& 11& 9 & 13& 7s \\
	17	& 13& 9 & 11& 6s \\
	18	& 9 & 13& 11& 6s \\
	19	& 13& 9 & 11& 5s \\
	20	& 9 & 11& 13& 5s \\
\end{dtable}

\paragraph{Features}

Choose or roll for a Feature roll for HP accordingly.

\begin{dtable}[cX|cX]
	\textbf{d10} & \textbf{Feature} & \textbf{d10} & \textbf{Feature} \\
	1 & Assassin	& 6 & Mystic \\
	2 & Berserker	& 7 & Skilled \\
	3 & Brawler		& 8 & Swashbuckler \\
	4 & Commander	& 9 & Tactician \\
	5 & Healer		& 10& Warrior \\
\end{dtable}

\subparagraph{Skilled} rolls for an expertise.

\begin{dtable}[cX|cX]
	\textbf{d8} & \textbf{Expertise} & \textbf{d8} & \textbf{Expertise} \\
	1 & animal handling	& 5 & drinking \\
	2 & athletics		& 6 & quickness \\
	3 & burgling		& 7 & stealth \\
	4 & cheating		& 8 & negotiation \\
\end{dtable}

\subparagraph{Mystic} rolls for random Cantrips (d20) and \nth{1} Circle Spells (``Random 36 Spells'' table from Chapter~4: Magic, p.17). Chose one of these spells as a Signature Spell or roll a d6 for a random one.

\paragraph{Background}

Chose or roll for a background and on any additional random table required.

\begin{dtable}[cX|cX]
	\textbf{d6} & \textbf{Background} & \textbf{d6} & \textbf{Background} \\
	1 & Criminal	& 4 & Sailor \\
	2 & Minstrel	& 5 & Scholar \\
	3 & Noble		& 6 & Soldier \\
\end{dtable}

\header{Criminal}
\begin{dtable}[cX|cX]
	\textbf{d6} & \textbf{Criminal Tool} & \textbf{d6} & \textbf{Criminal Tool} \\
	1 & blackjack (sap)	& 4 & loaded dice \\
	2 & crowbar			& 5 & lockpicks \\
	3 & grappling hook	& 6 & marked cards \\
\end{dtable}

\header{Minstrel}
\begin{dtable}[cX]
	\textbf{d10} & \textbf{Musical Instrument} \\
	1 & bagpipe, bladder pipe \\
	2 & drum, tambourine \\
	3 & flute, ocarina \\
	4 & jaw harp \\
	5 & harp, lyre \\
	6 & crumhorn, shawm \\
	7 & hurdy-gurdy \\
	8 & lute, mandolin \\
	9 & viol, fiddle, rebec \\
	10& zither, dulcimer \\
\end{dtable}

\header{Scholar}
\begin{dtable}[cX]
	\textbf{d12} & 	\textbf{Study} \\
	1	& History (archaeology) \\
	2	& History (culture and religion) \\
	3	& History (geography and politics) \\
	4	& Life Science (herbalism) \\
	5	& Life Science (medicine) \\
	6	& Life Science (zoology) \\
	7	& Philology (dead languages) \\
	8	& Philology (folklore and literature) \\
	9	& Philology (foreign languages) \\
	10	& Physical Science (astronomy and physics) \\
	11	& Physical Science (chemistry) \\
	12	& Physical Science (geoscience) \\
\end{dtable}


\header{Soldier}
\begin{dtable}[cXX]
	\textbf{d6} & \textbf{Rank} & \textbf{Weapon} \\
	1	& archer & longbow \\
	2	& cavalry & lance \\
	3	& musketeer & musket \\
	4	& officer & pistol \\
	5	& pikeman & pike \\
	6	& swordsman & claymore \\
\end{dtable}


\paragraph{Equipment}

Exchange your money for rolls on random equipment tables. Reroll duplicates. Mystics replace light armour with martial melee weapon.

\begin{dtable}[lX]
	\textbf{Money} & \textbf{Equipment} \\
	%4s	& simple M, 2$\times$gear, 2s \\
	5s	& simple M, 2$\times$gear, tool, 2s \\
	6s	& simple M, simple R, 2$\times$gear, tool, 2s \\
	7s	& simple M, shield, 1s \\
	8s	& simple M, simple R, pet, 1s \\
	%9s	& simple M, shield, 2$\times$gear, tool, 1s \\
	%10s	& simple M, simple R, pet, 2$\times$gear, tool, 1s \\
	10s	& simple M, shield, 2$\times$gear, tool, 2s \\
	%11s	& simple M, shield, pet \\
	12s	& simple M, military R, 1s \\
	14s	& simple M, light armour, 2$\times$gear, tool, 1s \\
	16s	& simple R., light armour, pet \\
	%18s	& simple M, military R, pet, 2$\times$gear, 1s \\
	%20s	& military M, shield, 2$\times$gear, tool, 3s \\
	%22s & military M, simple R, light armour, 1s \\
\end{dtable}

Additionally you have simple clothes. a backpack, basic camping equipment, torches, and a few days' rations.

\header{Simple Melee Weapons}
\begin{dtable}[cX|cX]
	\textbf{d4} & \textbf{Weapon} & \textbf{d4} & \textbf{Weapon} \\
	1 & pitchfork	& 3 & sledgehammer \\
	2 & quarterstaff& 4 & splitting maul \\
\end{dtable}

\header{Martial Melee Weapons}
\begin{dtable}[cX|cX]
	\textbf{d8} & \textbf{Weapon} & \textbf{d8} & \textbf{Weapon} \\
	1 & axe		& 5 & mace \\
	2 & dagger	& 6 & spear \\
	3 & halberd	& 7 & sword \\
	4 & lance	& 8 & war hammer \\
\end{dtable}

\vfill
\break

\header{Simple Ranged Weapons}
\begin{dtable}[cX|cX]
	\textbf{d6} & \textbf{Weapon} & \textbf{d6} & \textbf{Weapon} \\
	1 & boomerang	& 4 & hunting bow \\
	2 & darts		& 5 & sling \\
	3 & harpoon		& 6 & throwing daggers \\
\end{dtable}

\header{Martial Ranged Weapons}
\begin{dtable}[cX|cX]
	\textbf{d4} & \textbf{Weapon} & \textbf{d4} & \textbf{Weapon} \\
	1 & crossbow	& 3 & musket \\
	2 & longbow		& 4 & pistol \\
\end{dtable}

\header{Adventuring Gear}
\begin{dtable}[cX|cX]
	\textbf{d12} & \textbf{Gear} & \textbf{d12} & \textbf{Gear} \\
	1 & 10ft rope		& 7 & dice or cards \\
	2 & bottle			& 8 & flint and steel \\
	3 & caltrops		& 9	& parchment \\
	4 & candle			& 10& sack \\
	5 & chain			& 11& spike \\
	6 & chalk			& 12& tent \\
\end{dtable}

\header{Tools}
\begin{dtable}[cX|cX]
	\textbf{d20} & \textbf{Tool} & \textbf{d20} & \textbf{Tool} \\
	1 & animal trap			& 11& hatchet \\
	2 & collapsible pole	& 12& lockpicks \\
	3 & chisel and mallet	& 13& padlock \\
	4 & clamp				& 14& pickaxe \\
	5 & crowbar				& 15& pliers \\
	6 & drill				& 16& saw \\
	7 & file or rasp		& 17& scissors \\
	8 &	fishing pole		& 18& shovel \\
	9 & grappling hook		& 19& wrench \\
	10& hammer				& 20& writing set \\
\end{dtable}

\header{Pet}
\begin{dtable}[cX|cX]
	\textbf{d4} & \textbf{Pet} & \textbf{d4} & \textbf{Pet} \\
	1 & cat 	& 3 & owl \\
	2 & mutt	& 4 & parrot \\
\end{dtable}


\vfill
\break


\subsection{Random Magic Item Appearance}
\index{Magic Items}

Roll for a random magic item and it's appearance. Think of it's properties based on what it looks like.

\header{Type}
\begin{dtable}[cXcl]
	\textbf{d100} & \textbf{Type} & \textbf{d100} & \textbf{Type} \\
	1--10	&	container	&	71--73	&	music. instrument	\\
	11--30	&	expendable	&	74--80	&	light armour	\\
	31--40	&	garment	&	81--83	&	full armour	\\
	41--50	&	jewellery	&	84--90	&	shield	\\
	51--70	&	misc.	&	91--100	&	weapon	\\
\end{dtable}

\header{Container}
\begin{dtable}[cXcl]
	\textbf{d12} & \textbf{Container} & \textbf{d12} & \textbf{Container} \\
	1 & backpack or haversack	& 7 & jug \\
	2 & bottle					& 8 & pouch \\
	3 & box or casket			& 9 & quiver \\
	4 & decanter				& 10& sack or bag \\
	5 & drinking horn			& 11& vial \\
	6 & flask or canteen		& 12& waterskin \\
\end{dtable}

\header{Expendable}
\begin{dtable}[clcX]
	\textbf{d10} & \textbf{Expendable} & \textbf{d10} & \textbf{Expendable} \\
	1 & bean or seed				& 6 & herb \\
	2 & candle or torch				& 7 & ink \\
	3 & chalk						& 8 & oil \\
	4 & dust						& 9 & ointment \\
	5 & food (fruit, pastry, etc.)	& 10& potion \\
\end{dtable}

\header{Garment}
\begin{dtable}[cXcX]
	\textbf{d20} & \textbf{Garment} & \textbf{d20} & \textbf{Garment} \\
	1	& belt		& 11 & jerkin \\
	2	& boots		& 12 & mantle \\
	3	& cloak		& 13 & robe \\
	4	& coat		& 14 & sandals \\
	5	& doublet	& 15 & shirt \\
	6	& dress		& 16 & shoes \\
	7	& gloves	& 17 & skirt \\
	8	& hat		& 18 & trousers \\
	9	& hood		& 19 & tunic \\
	10	& hose		& 20 & vestments \\
\end{dtable}


\header{Jewellery}
\begin{dtable}[clcX]
	\textbf{d20} & \textbf{Jewellery} & \textbf{d20} & \textbf{Jewellery} \\
	1	& anklet 			& 11 & eyepatch \\
	2	& belt buckle 		& 12 & gorget \\
	3	& bracelet 			& 13 & hairpin \\
	4	& bracer 			& 14 & locket \\
	5	& brooch 			& 15 & mask \\
	6	& chain 			& 16 & medallion \\
	7	& cloak pin 		& 17 & necklace \\
	8	& crown 			& 18 & pectoral \\
	9	& diadem or tiara 	& 19 & pendant \\
	10	& earring 			& 20 & ring \\
\end{dtable}


\header{Miscellaneous}
\begin{dtable}[cXcl]
	\textbf{d100} & \textbf{Item} & \textbf{d100} & \textbf{Item} \\
	1--2	&	amulet or talisman	&	51--52	&	monocle	\\
	3--4	&	book	&	53--54	&	needle	\\
	5--6	&	bowl or bucket	&	55--56	&	orb or crystal	\\
	7--8	&	brazier	&	57--58	&	pickaxe	\\
	9--10	&	broom	&	59--60	&	pipe	\\
	11--12	&	brush	&	61--62	&	plate or tray	\\
	13--14	&	candelabrum	&	63--64	&	prosthesis	\\
	15--16	&	carpet	&	65--66	&	quill	\\
	17--18	&	censer	&	67--68	&	rod or sceptre	\\
	19--20	&	coin	&	69--70	&	rope	\\
	21--22	&	comb	&	71--72	&	saddle	\\
	23--24	&	corkscrew	&	73--74	&	sand timer	\\
	25--26	&	cup, chalice, or goblet	&	75--76	&	scissors	\\
	27--28	&	deck of cards or dice set	&	77--78	&	shovel	\\
	29--30	&	fan	&	79--80	&	sickle	\\
	31--32	&	figurine or idol	&	81--82	&	skull	\\
	33--34	&	gem or pearl	&	83--84	&	spectacles	\\
	35--36	&	hammer	&	85--86	&	spike	\\
	37--38	&	handkerchief	&	87--88	&	spyglass	\\
	39--40	&	horseshoe	&	89--90	&	staff	\\
	41--42	&	lantern	&	91--92	&	tablecloth	\\
	43--44	&	lens	&	93--94	&	tablet	\\
	45--46	&	lockpick	&	95--96	&	umbrella	\\
	47--48	&	manacles	&	97--98	&	wand	\\
	49--50	&	mirror	&	99--100	&	whistle	\\
\end{dtable}

\header{Musical Instrument}
\begin{dtable}[cXcX]
	\textbf{d20} & \textbf{Instrument} & \textbf{d20} & \textbf{Instrument} \\
	1	& bagpipe 		& 11 & jaw harp \\
	2	& bell 			& 12 & lute \\
	3	& bladder pipe	& 13 & lyre \\
	4	& crumhorn 		& 14 & mandolin \\
	5	& drum 			& 15 & ocarina \\
	6	& dulcimer 		& 16 & rebec \\
	7	& fiddle 		& 17 & shawm \\
	8	& flute 		& 18 & tambourine \\
	9	& harp 			& 19 & viol \\
	10	& hurdy-gurdy 	& 20 & zither \\
\end{dtable}

\header{Light Armour and Accessoires}
\begin{dtable}[cXcX]
	\textbf{d6} & \textbf{Armour} & \textbf{d6} & \textbf{Armour} \\
	1 & bracers		& 4 & greaves \\
	2 & gambeson	& 5 & helmet \\
	3 & gloves		& 6 & leather armour \\
\end{dtable}

\header{Full Armour and Accessoires}
\begin{dtable}[cXcl]
	\textbf{d10} & \textbf{Armour} & \textbf{d10} & \textbf{Armour} \\
	1 & bracers		& 6 & mail armour \\
	2 & cuirass		& 7 & plate armour \\
	3 & gauntlets	& 8 & sabatons \\
	4 & greaves		& 9 & scale armour \\
	5 & helmet		& 10& segmented armour \\
\end{dtable}

\header{Shield}
\begin{dtable}[cXcX]
	\textbf{d6} & \textbf{Shield} & \textbf{d6} & \textbf{Shield} \\
	1 & buckler			& 4 & pavise \\
	2 & heater shield	& 5 & round shield \\
	3 & kite shield		& 6 & square shield \\
\end{dtable}

\header{Weapon and Ammunition}
\begin{dtable}[cXcX]
	\textbf{d20} & \textbf{Weapon} & \textbf{d20} & \textbf{Weapon} \\
	1 & arrow		& 11 & hunting bow \\
	2 & axe			& 12 & lance \\
	3 & bolt		& 13 & longbow \\
	4 & boomerang	& 14 & pistol \\
	5 & bullet		& 15 & mace \\
	6 & crossbow	& 16 & musket \\
	7 & dagger		& 17 & sling \\
	8 & dart		& 18 & spear \\
	9 & halberd		& 19 & sword \\
	10& harpoon		& 20 & war hammer \\
\end{dtable}

\subsubsection{Random Appearance}

\header{Attribute}
\begin{dtable}[cXcX]
	\textbf{d20} & \textbf{Attribute} & \textbf{d20} & \textbf{Attribute} \\
	1	&	ancient	&	11	&	menacing	\\
	2	&	bejewelled	&	12	&	ornate	\\
	3	&	colourful	&	13	&	otherworldly	\\
	4	&	crude	&	14	&	patterned	\\
	5	&	dingy	&	15	&	peculiar	\\
	6	&	exotic	&	16	&	refined	\\
	7	&	grotesque	&	17	&	rugged	\\
	8	&	heavy	&	18	&	shiny	\\
	9	&	intricate	&	19	&	sleek	\\
	10	&	light	&	20	&	sophisticated	\\
\end{dtable}


\header{Fabric}
\begin{dtable}[cXcX]
	\textbf{d8} & \textbf{Fabric} & \textbf{d8} & \textbf{Fabric} \\
	1 & cotton	& 5 & leather \\
	2 & felt	& 6 & linen \\
	3 & fur		& 7 & silk \\
	4 & hair 	& 8	& wool \\
\end{dtable}


\header{Material}
\begin{dtable}[cXcX]
	\textbf{d20} & \textbf{Material} & \textbf{d20} & \textbf{Material} \\
	1 & amber	& 11 & iron \\
	2 & bone	& 12 & ivory \\
	3 & brass	& 13 & jade \\
	4 & bronze	& 14 & jet \\
	5 & ceramic	& 15 & obsidian \\
	6 & copper	& 16 & pewter \\
	7 & coral	& 17 & silver \\
	8 & crystal	& 18 & steel \\
	9 & glass	& 19 & stone \\
	10& gold	& 20 & wood \\
\end{dtable}

\vfill
\break


\end{document}
